
ANALYZING THE EFFECT OF VIOLENT VIDEO GAME PLAY ON MATE VALUE AND OTHER MATING RELATED VARIABLES
========================================================

General housekeeping
--------------------

First load the libraries that we are using
```{r}
library(ggplot2)
library(knitr)
library(QuantPsyc)
library(PerformanceAnalytics)
library(heplots)
library(plyr)
require(gridExtra)
```

Load the data (Complete_data_survey2.csv).
Note the as.is thing which leaves all the character vectors as strings and does not make them into factors.

```{r}
data = read.csv(file.choose(),  na.strings =".", header=T, sep=",")
```

Attach and check the headers of each column

```{r}
names(data)
```

Recode gender into numerical values

```{r}
data$Gender2 <- revalue(data$Gender, c("Male"="1", "Female"="2"))
data$Gender2 <- as.numeric(as.character(data$Gender2))
```

Recode violent/non-violent treatment into numerical values

```{r}
data$VorNV2 <- revalue(data$VorNV, c("V"="1", "NV"="2"))
data$VorNV2 <- as.numeric(as.character(data$VorNV2))
```

Ensuring that certain columns are treated as categorical

```{r}
data$Sexuality <- factor(data$Sexuality)
```


Delete Rows where participants were unable to load or play game, had issues with the survey loading, or failed an attention check (49 people). 

```{r}
data<-data[-which(data$Removedata == "Yes"), ]
```


Here's mean centring the data
```{r}
#data$mc_MV1Total <- scale(data$MV1Total, center = TRUE, scale = TRUE)
#data$mc_SOIRTOTAL <- scale(data$SOIRTOTAL, center = TRUE, scale = TRUE)
#data$mc_MV2Average <- scale(data$MV2Average, center = TRUE, scale = TRUE)
#data$mc_SGAAverage <- scale(data$SGAAverage, center = TRUE, scale = TRUE)
#data$mc_HostilityTotal <- scale(data$HostilityTotal, center = TRUE, scale = TRUE)
#data$mc_MMTotal <- scale(data$MMTotal, center = TRUE, scale = TRUE)
```

---------------
First, check the distributions of all the variables of interest
```{r}
hist(data$MV1Total)
hist(data$MV2Average)
hist(data$SGAAverage)
hist(data$HostilityTotal)
hist(data$SOIRTotal)
hist(data$MMTotal)
hist(data$Matevalue)
```

The ditributions for all of them are fine, except for Hostility Total and Mating Motivation. HT specifically has very little variation in it, so I don't think it will be very useful and am happy to remove it from the models. MM has a terrible skew and no transformations will really help. As a result, I think it would be alright to remove that from the model as well.

What is the mean, SD and range of the age of respondents in the first sample?
```{r}
mean(data$Age)
sd(data$Age)
range(data$Age)
```


Do men and women differ in the amount of violent video games they play (self-rated)

```{r}
VVGSex1 <- lm(Averagefrequencyviolence ~ Age + Gender + Age*Gender, data = data)
anova(VVGSex1, test="Chisq")
summary(VVGSex1, test="Chisq")
```
HISTOGRAM
```{r}
hist1 <- ggplot(data, aes(x=Averagefrequencyviolence, fill=Gender)) + 
  geom_histogram(binwidth=2, position="dodge") +
  xlab("Self-rated Violence Exposure") +
  ylab("Count") +
  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
  theme(axis.line = element_line(colour = "black"),
        plot.title = element_text(size=30, face="bold"),
        legend.position = "right",
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 

hist1
```
Boxplot of VVGplay between gender (Self-reported)

```{r}
BoxVVGgender <- ggplot(data, aes(x=Gender, y=Averagefrequencyviolence)) + geom_boxplot() + 
  guides(fill=FALSE) + 
  xlab("Gender") +
  ylab("Self-Rated Violence Exposure") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 

BoxVVGgender 
```


Relationship between Age and Exposure to violence:
```{r}
plotAgeExposure <- ggplot(data, aes(x = Age, y = Averagefrequencyviolence, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(0,50) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(10,75)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Age") +
  ylab("Violent Video Game Exposure") +
    theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotAgeExposure
```


---------------------------

##MANOVA
```{r}
MANOVA2<- manova(cbind(Matevalue, SGAAverage, SOIRTotal, MMTotal) ~ VorNV + Age + Selfscore + Gender + Averagefrequencyviolence + VorNV*Age + VorNV*Selfscore + VorNV*Gender + Selfscore*Gender + Age*Selfscore, data=data)
summary(MANOVA2, test="Pillai")
```

##Univariate Tests
```{r}
summary.aov(MANOVA2)

lm1<- lm(Matevalue ~ VorNV + Age + Selfscore + Gender + Averagefrequencyviolence + VorNV*Age + VorNV*Selfscore + VorNV*Gender + Selfscore*Gender + Age*Selfscore, data=data)
summary(lm1)
anova(lm1)
anova(lm1, test = "Chisq")

plot(fitted(lm1), residuals(lm1))
qqnorm(residuals(lm1))
qqline(residuals(lm1))

lm2<- lm(Matevalue ~ VorNV + Age + Selfscore + Gender + Averagefrequencyviolence + VorNV*Age + VorNV*Selfscore + VorNV*Gender + Selfscore*Gender + Age*Selfscore , data=data)
summary(lm2)
anova(lm2, test = "Chisq")

plot(fitted(lm2), residuals(lm2))
qqnorm(residuals(lm2))
qqline(residuals(lm2))

lm3<- lm(SGAAverage ~ VorNV + Age + Selfscore + Gender + Averagefrequencyviolence + VorNV*Age + VorNV*Selfscore + VorNV*Gender + Selfscore*Gender + Age*Selfscore, data=data)
summary(lm3)
anova(lm3, test = "Chisq")

plot(fitted(lm3), residuals(lm3))
qqnorm(residuals(lm3))
qqline(residuals(lm3))

lm4<- lm(SOIRTotal ~ VorNV + Age + Selfscore + Gender + Averagefrequencyviolence + VorNV*Age + VorNV*Selfscore + VorNV*Gender + Selfscore*Gender + Age*Selfscore, data=data)
summary(lm4)
anova(lm4, test = "Chisq")

plot(fitted(lm4), residuals(lm4))
qqnorm(residuals(lm4))
qqline(residuals(lm4))

lm5<- lm(MMTotal ~ VorNV + Age + Selfscore + Gender + Averagefrequencyviolence + VorNV*Age + VorNV*Selfscore + VorNV*Gender + Selfscore*Gender + Age*Selfscore, data=data)
summary(lm5)
anova(lm5, test="Chisq")

plot(fitted(lm5), residuals(lm5))
qqnorm(residuals(lm5))
qqline(residuals(lm5))
```
---------------------------------------------------------

Sex Goal Activation (SGA)

SGA and Selfscore by Gender plot
```{r}
plotsgaage <- ggplot(data, aes(x = SGAAverage, y = Selfscore, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(0.5,7.5) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,7.5)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Sex Goal Activation") +
  ylab("Self Rated Score") +
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotsgaage
```
-------------------------------------------------------------
Short term faces: 

STselection <- lm(TotalSTScore ~ MV2Average + SGAAverage + SOIRTotal + Age + Gender + Gender*MV2Average + Gender*SGAAverage + Gender*SOIRTotal, data=data)
anova(STselection)


```{r}
plotstscore <- ggplot(data, aes(x = Selfscore, y = TotalSTScore, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(0,7.5) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,7.5)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Facial Preference for masculinity ") +
  ylab("Self Rated Score") + 
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotstscore
```


LT Faces and age
```{r}
plotltscore <- ggplot(data, aes(x = TotalLTScore, y = Age, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(18,76) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,7.5)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Facial Preference for masculinity ") +
  ylab("Age") + 
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotltscore
```
-----

SOIR (Revised Sociosexual Orientation Inventory) 

SOIR + Averagefrequencyviolence+Gender Regression
```{r}
fitSOI <- lm(SOIRTotal ~ Averagefrequencyviolence + Age+ Gender, data=data)
summary(fitSOI)
```


SOIR Age and Gender Plot
```{r}
plotSOI1 <- ggplot(data, aes(x = Age, y = SOIRTotal, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size=4) +
  ylim(5,80) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(15,80)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Age") +
  ylab("Sexual Interest") +
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotSOI1
```

SOIR V or NV gender
```{r}
#plotSOI2 <- ggplot(data, aes(x = VorNV, y = SOIRTotal, color = Gender, shape = Gender, linetype = Gender)) + 
#  geom_point(size=4) +
#  ylim(5,80) +
#  scale_colour_hue(l=50, guide = FALSE) +
#  scale_x_continuous(limits=c(15,80)) +
#  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
#  xlab("Treatment") +
#  ylab("Sexual Interest") +
#  theme(legend.position="right") +
#  theme(axis.line = element_line(colour = "black"),
#        axis.text=element_text(size=18),
#        axis.title=element_text(size=20,face="bold"),
#        axis.title.y=element_text(vjust=0.25),
#        axis.title.x=element_text(vjust=0.25),
#        panel.grid.major = element_blank(),
#        panel.grid.minor = element_blank(),
#        panel.border = element_blank(),
#        panel.background = element_blank()) 
#plotSOI2
```
-------------------------------------------------------------------

Mate Value Test 1 (4-Item Mate Value Scale, Edlund and Sagarin 2014)

MV1 + gender + selfscore ANOVA*
```{r}
MV1selfscore <- lm(MV1Total ~  Gender+ Selfscore + Selfscore*Gender, data = data)
anova(MV1selfscore, test="Chisq")
```

MV1 x selfscore x gender plot 
```{r}
plotMv1selfs <-ggplot(data, aes(x = Selfscore, y = MV2Average, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0.5,7.5), breaks=c(1:7)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Selfscore") +
  ylab("Mate Value") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotMv1selfs
```
MV1 x TotalSTScore x gender plot 
```{r}
plotMv1selfs <-ggplot(data, aes(x = TotalSTScore, y = MV1Total, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0.5,7.5), breaks=c(1:7)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Selfscore") +
  ylab("Mate Value") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotMv1selfs
```

Mating motivation: 

Matingmotivation + gender + selfscore + Violence exposure ANOVA 

```{r}
MMANOVAVVG <- lm(MMTotal ~  Gender + Age + Gender*Averagefrequencyviolence + Averagefrequencyviolence + Selfscore + Selfscore*Gender, data = data)
anova(MMANOVAVVG, test="Chisq")
```

Plot Matingmoviation and Violence Exposure
```{r}
ggplot(data, aes(x = SOIRTotal, y = TotalSTScore, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size=4) +
  scale_colour_hue(l=50, guide = FALSE) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("SOIRTotal") +
  ylab("Facial Preference") +
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 

```


MATEVALUE

```{r}
plotmvselfscore <- ggplot(data, aes(x = Matevalue, y = Selfscore, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(0.5,7.5) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,7.5)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Mate value ") +
  ylab("Self Rated Score") + 
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotmvselfscore
```
MV + STScore
```{r}
plotstsmv <- ggplot(data, aes(x = TotalSTScore, y = Matevalue, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(0.5,7.5) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,7.5)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Facial Preference for masculinity ") +
  ylab("MV") + 
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotstsmv
```

SOIR And STScore
```{r}
SOIRSTSAnova <- lm(TotalSTScore ~ Gender + Age + SOIRTotal + Gender*SOIRTotal, data = data)
anova(SOIRSTSAnova, test="Chisq")
```

SOIR and LTScore 
```{r}
SOIRLTSAnova <- lm(TotalLTScore ~ Gender + Age + SOIRTotal + Gender*SOIRTotal, data = data)
anova(SOIRLTSAnova, test="Chisq")
```

LTScore + SOIR
```{r}
plotltssoir <- ggplot(data, aes(x = TotalLTScore, y = SOIRTotal, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(5,80) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,7.5)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Facial Preference for masculinity ") +
  ylab("SOIR") + 
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotltssoir
```
MV + STScore
```{r}
plotstsmv <- ggplot(data, aes(x = TotalSTScore, y = Matevalue, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(0.5,7.5) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,7.5)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Facial Preference for masculinity ") +
  ylab("MV") + 
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotstsmv
```