
ANALYZING THE EFFECT OF VIOLENT VIDEO GAME PLAY ON MATE VALUE AND OTHER MATING RELATED VARIABLES
========================================================

General housekeeping
--------------------

First load the libraries that we are using
```{r}
library(ggplot2)
library(knitr)
library(QuantPsyc)
library(PerformanceAnalytics)
library(heplots)
library(plyr)
require(gridExtra)
```

Load the data (Complete_data_survey.csv).
Note the as.is thing which leaves all the character vectors as strings and does not make them into factors.

```{r}
data = read.csv(file.choose(),  na.strings =".", header=T, sep=",")
```

Attach and check the headers of each column

```{r}
names(data)
```

Recode gender into numerical values

```{r}
#data$Gender2 <- revalue(data$Gender, c("Male"="1", "Female"="2"))
#data$Gender2 <- as.numeric(as.character(data$Gender2))
data$Numericalgameplayed <- factor(data$Numericalgameplayed)
```
Recode violent/non-violent treatment into numerical values

```{r}
data$VorNV2 <- revalue(data$VorNV, c("V"="1", "NV"="2"))
data$VorNV2 <- as.numeric(as.character(data$VorNV2))
```
Ensuring that certain columns are treated as categorical

```{r}
data$Sexuality <- factor(data$Sexuality)

```


Delete Rows where participants were unable to load or play game, had issues with the survey loading, or failed an attention check (27 people). Also delete rows where participants recorded gender as "Other" (5 people).

```{r}
data<-data[-which(data$Removedata == "Yes"), ]
data<-data[-which(data$Gender == "Other"), ]
```


Remove individuals that have a relationship status of 'other' (5 individuals)

```{r}
data = subset(data, Relationshipstatus %in% c("Single", "In an open relationship/casually dating", "In a long-term monogamous relationship eg. married, partnered", "Recently single/divorced/separated"))
```

Here's mean centring the data
```{r}
#data$mc_MV1Total <- scale(data$MV1Total, center = TRUE, scale = TRUE)
#data$mc_SOIRTOTAL <- scale(data$SOIRTOTAL, center = TRUE, scale = TRUE)
#data$mc_MV2Average <- scale(data$MV2Average, center = TRUE, scale = TRUE)
#data$mc_SGAAverage <- scale(data$SGAAverage, center = TRUE, scale = TRUE)
#data$mc_HostilityTotal <- scale(data$HostilityTotal, center = TRUE, scale = TRUE)
#data$mc_MMTotal <- scale(data$MMTotal, center = TRUE, scale = TRUE)
```


---------------
First, check the distributions of all the variables of interest
```{r}
hist(data$MV1Total)
hist(data$MV2Average)
hist(data$SGAAverage)
hist(data$HostilityTotal)
hist(data$SOIRTOTAL)
hist(data$MATINGMOTIVATIONtotal)
```

The ditributions for all of them are fine, except for Hostility Total and Mating Motivation. HT specifically has very little variation in it, so I don't think it will be very useful and am happy to remove it from the models. MM has a terrible skew and no transformations will really help. As a result, I think it would be alright to remove that from the model as well.

What is the mean, SD and range of the age of respondents in the first sample?
```{r}
mean(data$Age)
sd(data$Age)
range(data$Age)
```

How many men and women, and their relationship details
```{r}
table(data$Gender)
table(data$Relationshipstatus)
table(data$lengthofrelationship)
count(data, c("Kinsey", "Gender"))
count(data, c("Relationshipstatus", "Gender"))
```

Do men and women differ in the amount of violent video games they play (self-rated)

```{r}
VVGDemo <- lm(Averagefrequencyviolence ~  Age + Kinsey + Gender + Age*Kinsey + Gender*Kinsey + Age*Gender, data = data)
anova(VVGDemo, test="Chisq")
summary(VVGDemo, test="Chisq")

```

HISTOGRAM
```{r}
ggplot(data, aes(x=Averagefrequencyviolence, fill=Gender)) + 
  geom_histogram(binwidth=2, position="dodge") +
  xlab("Self-rated Violence Exposure") +
  ylab("Count") +
  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
  theme(axis.line = element_line(colour = "black"),
        plot.title = element_text(size=30, face="bold"),
        legend.position = "right",
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
```

Relationship between Kinsey Scale, Gender, and Exposure to violence:
```{r}
ggplot(data, aes(x = Kinsey, y = Averagefrequencyviolence, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(0,50) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,7)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Kinsey Scale") +
  ylab("Violent Video Game Exposure") +
    theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
```

Relationship between Age and Exposure to violence:
```{r}
ggplot(data, aes(x = Age, y = Averagefrequencyviolence, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(0,50) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,7)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Age") +
  ylab("Violent Video Game Exposure") +
    theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
```


---------------------------



Do men and women differ in how well they performed in the game in both self-rated (selfscore) and actual performance

```{r}
Perf <- lm(Selfscore ~  Age + Gender + VorNV + Age*VorNV + Gender*VorNV, data = data)
anova(Perf, test="Chisq")
summary(Perf, test="Chisq")


ggplot(data, aes(x = Age, y = Selfscore, color = VorNV, shape = VorNV, linetype = VorNV)) + 
  geom_point(size = 4) + 
  ylim(0,10) +
  scale_colour_hue(l=50, guide = FALSE) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Age") +
  ylab("Selfscore") +
    theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 

ggplot(data, aes(x = Age, y = Selfscore, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(0,10) +
  scale_colour_hue(l=50, guide = FALSE) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Age") +
  ylab("Selfscore") +
    theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 

##Need to find some way to standardize the scores within game category
##Nevermind...people lied with their scores so standardizing doesn't make a difference
#ddply(data, "Gameplayed", transform, Score_std = scale(data$Score, center = TRUE, scale = TRUE))
```

##MANOVA
```{r}
MANOVA1<- manova(cbind(MV2Average, SGAAverage, SOIRTOTAL, MATINGMOTIVATIONtotal) ~ VorNV + Age + Selfscore + Kinsey + Gender + Averagefrequencyviolence + VorNV*Selfscore + VorNV*Gender, data=data)
summary(MANOVA1, test="Pillai")
```

##Univariate Tests
```{r}
summary.aov(MANOVA1)

lm1<- lm(MV1Total ~ VorNV + Age + Selfscore + Gender + Averagefrequencyviolence + VorNV*Age + VorNV*Selfscore + VorNV*Gender + Selfscore*Gender + Age*Selfscore, data=data)
summary(lm1)
anova(lm1)
anova(lm1, test = "Chisq")

plot(fitted(lm1), residuals(lm1))
qqnorm(residuals(lm1))
qqline(residuals(lm1))

lm2<- lm(MV2Average ~ VorNV + Age + Selfscore + Gender + Averagefrequencyviolence + VorNV*Age + VorNV*Selfscore + VorNV*Gender + Selfscore*Gender + Age*Selfscore, data=data)
summary(lm2)
anova(lm2, test = "Chisq")

plot(fitted(lm2), residuals(lm2))
qqnorm(residuals(lm2))
qqline(residuals(lm2))

lm3<- lm(SGAAverage ~ VorNV + Age + Selfscore + Gender + Averagefrequencyviolence + VorNV*Age + VorNV*Selfscore + VorNV*Gender + Selfscore*Gender + Age*Selfscore, data=data)
summary(lm3)
anova(lm3, test = "Chisq")

plot(fitted(lm3), residuals(lm3))
qqnorm(residuals(lm3))
qqline(residuals(lm3))

lm4<- lm(SOIRTOTAL ~ VorNV + Age + Selfscore + Gender + Averagefrequencyviolence + VorNV*Age + VorNV*Selfscore + VorNV*Gender + Selfscore*Gender + Age*Selfscore, data=data)
summary(lm4)
anova(lm4, test = "Chisq")

plot(fitted(lm4), residuals(lm4))
qqnorm(residuals(lm4))
qqline(residuals(lm4))

lm5<- lm(MATINGMOTIVATIONtotal ~ VorNV + Age + Selfscore + Gender + Averagefrequencyviolence + VorNV*Age + VorNV*Selfscore + VorNV*Gender + Selfscore*Gender + Age*Selfscore, data=data)
summary(lm5)
anova(lm5, test="Chisq")

plot(fitted(lm5), residuals(lm5))
qqnorm(residuals(lm5))
qqline(residuals(lm5))
```
---------------------------------------------------------
Mate Value Test 2 (Mate Value Inventory - Kirsner et al 2003)

MV2 + Age + Selfscore*gender ANOVA
```{r}
VorNVMV2 <- lm(MV2Average ~ Age + Gender + Selfscore + Selfscore*Gender + Selfscore*Age, data = data)
anova(VorNVMV2, test="Chisq")
```

Plot MV 2 and selfscore : 

```{r}
plotMvAverageselfs <- ggplot(data, aes(x = Selfscore, y = MV2Average, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(0,7.5) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0.5,7.5), breaks=c(1:7)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Selfscore") +
  ylab("Mate Value") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotMvAverageselfs
```

Plot MV2 and age 
```{r}
plotmv2age <- ggplot(data, aes(x = Age, y = MV2Average, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(1,7) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(10,70)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Age") +
  ylab("Mate Value") +
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotmv2age
```
--------------------------------------------------------
Sex Goal Activation (SGA)


SGA and AGE Plot
```{r}
plotsgaage <- ggplot(data, aes(x = Age, y = SGAAverage, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(0.5,7.5) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(10,70)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Age") +
  ylab("Sex Goal Activation") +
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotsgaage
```
-------------------------------------------------------------
SOIR (Revised Sociosexual Orientation Inventory) 

SOIR + Averagefrequencyviolence+Gender Regression
```{r}
fitSOI <- lm(SOIRTOTAL ~ Averagefrequencyviolence + Gender, data=data)
summary(fitSOI)
```
Plot of exposure to violence on SOIR 
```{r}
plotSOI1 <- ggplot(data, aes(x = SOIRTOTAL, y = Averagefrequencyviolence, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size=4) +
  ylim(0,50) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,80)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Sexual Interest") +
  ylab("Violent Video Game Exposure") +
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotSOI1
```


SOIR Age and Gender Plot
```{r}
plotSOI2 <- ggplot(data, aes(x = Age, y = SOIRTOTAL, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size=4) +
  ylim(5,75) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(15,80)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Age") +
  ylab("Sexual Interest") +
  theme(legend.position="none") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotSOI2
```

SOIR Self Score plot
```{r}
plotSOI3 <- ggplot(data, aes(x = Selfscore, y = SOIRTOTAL, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size=4) +
  ylim(7,75) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0.5,9.5)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Self-Rated Score") +
  ylab("Sexual Interest") +
  theme(legend.position="none") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotSOI3
```

-------------------------------------------------------------------

MATE VALUE TEST 1 AND MATING MOTIVATION TESTS : (TO REMOVE)
Mate Value Test 1 (4-Item Mate Value Scale, Edlund and Sagarin 2014)

MV1 + gender + selfscore ANOVA*
```{r}
MV1selfscore <- lm(MV1Total ~  Gender+ Selfscore + Selfscore*Gender, data = data)
anova(MV1selfscore, test="Chisq")
```

MV1 x selfscore x gender plot 
```{r}
plotMv1selfs <-ggplot(data, aes(x = Selfscore, y = MV1Total, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0.5,9), breaks=c(1:7)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Selfscore") +
  ylab("Mate Value") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotMv1selfs
```

Mating motivation: 

Matingmotivation + gender + selfscore + Violence exposure ANOVA 

```{r}
MMANOVAVVG <- lm(MATINGMOTIVATIONtotal ~  Gender + Gender*Averagefrequencyviolence + Averagefrequencyviolence + Selfscore + Selfscore*Gender, data = data)
anova(MMANOVAVVG, test="Chisq")
```

Plot Matingmoviation and selfscore
```{r}
plotmmselfscore <- ggplot(data, aes(x = Selfscore, y = MATINGMOTIVATIONtotal, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size=4) +
  ylim(0,30) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,7.5)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Self Score") +
  ylab("Gaming Motivation") +
  theme(legend.position="right") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotmmselfscore
```
