ANALYZING THE EFFECT OF VIOLENT VIDEO GAME PLAY ON MATE VALUE AND OTHER MATING RELATED VARIABLES
========================================================

General housekeeping
--------------------

First load the libraries that we are using
```{r}
library(ggplot2)
library(knitr)
library(QuantPsyc)
library(PerformanceAnalytics)
library(heplots)
library(plyr)
require(gridExtra)
```

Next load the functions that we'll use for summary data

Using the script written here: http://www.r-bloggers.com/more-on-exploring-correlations-in-r/
to create the correlation data for the table, but also to create a figure of correlation values

See http://goo.gl/nahmV for documentation of this function
```{r}
cor.prob <- function (X, dfr = nrow(X) - 2) {
  R <- cor(X, use="pairwise.complete.obs")
  above <- row(R) < col(R)
  r2 <- R[above]^2
  Fstat <- r2 * dfr/(1 - r2)
  R[above] <- 1 - pf(Fstat, 1, dfr)
  R[row(R) == col(R)] <- NA
  R
}
```

Use this function to dump the cor.prob output to a 4 column matrix with row/column indices, correlation, and p-value. See StackOverflow question: http://goo.gl/fCUcQ

```{r}
flattenSquareMatrix <- function(m) {
  if( (class(m) != "matrix") | (nrow(m) != ncol(m))) stop("Must be a square matrix.") 
  if(!identical(rownames(m), colnames(m))) stop("Row and column names must be equal.")
  ut <- upper.tri(m)
  data.frame(i = rownames(m)[row(m)[ut]],
             j = rownames(m)[col(m)[ut]],
             cor=t(m)[ut],
             p=m[ut])
}
```

Use this one for summary statistics

```{r}
my.summary <- function(x, na.rm=TRUE){
  result <- c(Mean=mean(x, na.rm=na.rm),
              SD=sd(x, na.rm=na.rm),
              Median=median(x, na.rm=na.rm),
              Min=min(x, na.rm=na.rm),
              Max=max(x, na.rm=na.rm), 
              N=length(x))
}
```

Load the data (Complete_data_survey.csv).
Note the as.is thing which leaves all the character vectors as strings and does not make them into factors.

```{r}
data = read.table(file.choose(),  na.strings =".", header=T, sep=",")
```

Attach and check the headers of each column

```{r}
attach(data)
names(data)
```

Recode gender into numerical values

```{r}
data$Gender2 <- revalue(data$sex_response, c("Male"="1", "Female"="2"))
data$Gender2 <- as.numeric(as.character(data$Gender2))
```

Ensuring that certain columns are treated as categorical

```{r}
data$sexuality_response <- factor(data$sexuality_response)
data$age_response <- factor(data$age_response)
data$soirq1_response <- factor(data$soirq1_response)
data$soirq2_response <- factor(data$soirq2_response)
data$soirq3_response <- factor (data$soirq3_response)
```


Delete Rows where participants were unable to load or play game, or remember score

```{r}
data<-data[-which(is.na(data$Removedata)),]
```

Then subset data by Study

```{r}
mturk1 = subset(data, Study=="Mturk1")
```

Remove individuals that have a relationship status of 'other' (2 individuals)

```{r}
mturk1 = subset(mturk1, relationshipstatus_response %in% c("Single", "In an open relationship/casually dating", "In a long-term monogamous relationship eg. married, partnered", "Recently single/divorced/separated"))
```

Here's mean centring the data
```{r}
mturk1$mc_mvscaletotal <- scale(mturk1$mvscaletotal, center = TRUE, scale = TRUE)
mturk1$mc_SOIRTOTAL <- scale(mturk1$SOIRTOTAL, center = TRUE, scale = TRUE)
mturk1$mc_MvAverage <- scale(mturk1$MvAverage, center = TRUE, scale = TRUE)
mturk1$mc_SGAAverage <- scale(mturk1$SGAAverage, center = TRUE, scale = TRUE)
mturk1$mc_HostilityTotal <- scale(mturk1$HostilityTotal, center = TRUE, scale = TRUE)
mturk1$mc_MATINGMOTIVATION_total <- scale(mturk1$MATINGMOTIVATION_total, center = TRUE, scale = TRUE)
```

---------------

What is the mean, SD and range of the age of respondents in the first sample?
```{r}
mean(mturk1$age_response)
sd(mturk1$age_response)
range(mturk1$age_response)
```


Do men and women differ in the amount of violent video games they play (self-rated)

```{r}
VVGsexM1 <- lm(AverageFrequencyviolence_mean ~ sex_response + age_response + sex_response*age_response, data = mturk1)
anova(VVGsexM1, test="Chisq")
```

Plot to see it. Boxplot of VVGplay between gender (Self-reported)

```{r}
ggplot(mturk1, aes(x=sex_response, y=AverageFrequencyviolence_mean)) + geom_boxplot() + guides(fill=FALSE) + 
  xlab("Gender") +
  ylab("Self-Rated Violence Exposure") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
```

Yes they do; women play less violent games. 
Here is the correlation between self- and objective-rated violence in video games


Now make a histogram of the overlap in violent video game play between the sexes
```{r}
hist1 <- ggplot(mturk1, aes(x=AverageFrequencyviolence_mean, fill=sex_response)) + 
  geom_histogram(binwidth=2, position="dodge") +
  xlab("Self-rated Violence Exposure") +
  ylab("Count") +
  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
  theme(axis.line = element_line(colour = "black"),
        plot.title = element_text(size=30, face="bold"),
        legend.position = "none",
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 

hist1
```


```{r}
cor(mturk1$Averagefrequencyviolence_mean,mturk1$PlayXESRB_mean)
```

get the columns that you want from the dataset
then create a correlation matrix
then get the p-values using the above function
"flatten"" theat table using the above function

```{r}
corrmturk1 <- mturk1[, c(13,39,12,8,10,38,5,7,9,13)]
cor(corrmturk1)
cor.prob(corrmturk1)
flattenSquareMatrix(cor.prob(corrmturk1))
```

Finally, plot the data using, but turn off warnings first
```{r}
options(warn=-1)
chart.Correlation(corrmturk1)
```

Now to get the meands and SDs for the above selected columns using my.summary function

```{r}
sapply(corrmturk1, my.summary)  
```

and then the numbers in each of the categorical groups

```{r}
table(mturk1$Education)
table(mturk1$Sexuality)
table(mturk1$RelStat)
```

relevel the coefficients
```{r}
mturk1$RelStat <-relevel(mturk1$RelStat, "2")
mturk1$Sexuality <-relevel(mturk1$Sexuality, ref = "3")
mturk1$Education <-relevel(mturk1$Education, ref = "1")
```

Here is the full model

```{r}
VGplayself.all.mturk1<-lm(PlayXViolence_mean ~ MV1.new + SOI + LeadershipDom + WordSum + CTSlog + Sexuality + RelStat + Income + Education + Gender + Age + Gender*MV1.new + Gender*SOI + Gender*LeadershipDom + Gender*WordSum + Gender*CTSlog + Gender*Sexuality + Gender*RelStat + Gender*Income + Gender*Education, data=mturk1)

anova(VGplayself.all.mturk1, test="typeIII")
summary(VGplayself.all.mturk1)

etasq(VGplayself.all.mturk1)
```

calculate r-squared for relationship between SOI, Education & MV with VVGExposure

```{r}
r_mturk1_SOI <- lm(PlayXViolence_mean ~ SOI, data=mturk1)
summary(r_mturk1_SOI)

r_mturk1_EDU <- lm(PlayXViolence_mean ~ Education, data=mturk1)
summary(r_mturk1_EDU)

r_mturk1_MV.m <- lm(PlayXViolence_mean ~ MV1.new, data=subset(mturk1, Gender=="Male"))
summary(r_mturk1_MV.m)

r_mturk1_MV.f <- lm(PlayXViolence_mean ~ MV1.new, data=subset(mturk1, Gender=="Female"))
summary(r_mturk1_MV.f)

```

these are standardized coefficients...apparantly

```{r}
lm.beta(VGplayself.all.mturk1)
```

It seems SOI has an overall effect and there is an interaction between Mate Value and Gender

Plot Mate Value and VVGplay for each gender

```{r}
plotMV1 <- ggplot(mturk1, aes(x = MV1.new, y = PlayXViolence_mean, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size = 4) + 
  ylim(0,50) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0.5,7.5), breaks=c(1:7)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Mate Value") +
  ylab("Violent Video Game Exposure") +
  theme(legend.position="none") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotMV1
```

Now plot SOI
```{r}
plotSOI1 <- ggplot(mturk1, aes(x = SOI, y = PlayXViolence_mean, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size=4) +
  ylim(0,50) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,80)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Sexual Interest") +
  ylab("Violent Video Game Exposure") +
  theme(legend.position="none") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
plotSOI1
```

Boxplot of VVGplay with education  (Self-reported)

```{r}
BoxEDU1 <- ggplot(mturk1, aes(x=Education, y=PlayXViolence_mean)) + geom_boxplot() + 
  guides(fill=FALSE) + 
  xlab("Level of Education") +
  ylab("Self-Rated Violence Exposure") +
  scale_x_discrete(labels=c("Primary", "High School", "Vocational", "University", "Post-Graduate"))+
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 

BoxEDU1
```

Explore 'MTurk2'
----------------

Do men and women differ in the amount of violent video games they play (self-rated)

```{r}
VVGsexM2 <- lm(PlayXViolence_mean ~ Gender + Age + Gender*Age, data = mturk2)
anova(VVGsexM2, test="Chisq")
```

Plot to see it. Boxplot of VVGplay between gender (Self-reported)

```{r}
ggplot(mturk2, aes(x=Gender, y=PlayXViolence_mean)) + geom_boxplot() + guides(fill=FALSE) + 
  xlab("Gender") +
  ylab("Self-Rated Violence Exposure") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
```

Yes they do; women play less violent games. 

Now make a histogram of the overlap in violent video game play between the sexes
```{r}
hist2 <- ggplot(mturk2, aes(x=PlayXViolence_mean, fill=Gender)) + 
  geom_histogram(binwidth=2, position="dodge") +
  xlab("Self-rated Violence Exposure") +
  ylab("Count") +
  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
  theme(axis.line = element_line(colour = "black"),
        plot.title = element_text(size=30, face="bold"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position=c(1,1),legend.justification=c(1,1)) 

hist2
```


Again, here is the coorrelation between self- and objective-ratings of violence

```{r}
cor(mturk2$PlayXViolence_mean,mturk2$PlayXESRB_mean)
```

Again, use the same function as above to get a table and figure of correlation values

```{r}
corrmturk2 <- mturk2[, c(13,6,7,8,10,39,12,17)]
cor(corrmturk2)
cor.prob(corrmturk2)
flattenSquareMatrix(cor.prob(corrmturk2))
```

Finally, plot the data
```{r}
chart.Correlation(corrmturk2)
```

Now to get the means and SDs for the above selected columns using my.summary function

```{r}
sapply(corrmturk2, my.summary)  
```

and then the numbers in each of the categorical groups

```{r}
table(mturk2$Education)
table(mturk2$Sexuality)
table(mturk2$RelStat)
```

relevel the coefficients
```{r}
mturk2$RelStat <-relevel(mturk2$RelStat, "2")
mturk2$Sexuality <-relevel(mturk2$Sexuality, ref = "3")
mturk2$Education <-relevel(mturk2$Education, ref = "1")
```

Here is the full model WITHOUT game motivation

```{r}
VGplayself.all1.mturk2<-lm(PlayXViolence_mean ~ Mvall + SOI + LeadershipDom + WordSum + Sexuality + RelStat + Income + Education + Gender + Age + Gender*Mvall + Gender*SOI + Gender*LeadershipDom + Gender*WordSum + Gender*Sexuality + Gender*RelStat + Gender*Income + Gender*Education, data=mturk2)

anova(VGplayself.all1.mturk2, test="Chisq")
summary(VGplayself.all1.mturk2)

etasq(VGplayself.all1.mturk2)
```

calculate r-squared for relationship between SOI, Education & MV with VVGExposure

```{r}
r_mturk2_SOI <- lm(PlayXViolence_mean ~ SOI, data=mturk2)
summary(r_mturk2_SOI)

r_mturk2_EDU <- lm(PlayXViolence_mean ~ Education, data=mturk2)
summary(r_mturk2_EDU)

r_mturk2_MV.m <- lm(PlayXViolence_mean ~ MV1.new, data=subset(mturk2, Gender=="Male"))
summary(r_mturk2_MV.m)

r_mturk2_MV.f <- lm(PlayXViolence_mean ~ MV1.new, data=subset(mturk2, Gender=="Female"))
summary(r_mturk2_MV.f)

r_mturk2_age <- lm(PlayXViolence_mean ~ Age, data=mturk2)
summary(r_mturk2_age)
```

Here is the full model WITH game motivation

```{r}
VGplayself.all2.mturk2<-lm(PlayXViolence_mean ~ Mvall + SOI + LeadershipDom + WordSum + Sexuality + RelStat + Income + Education + log(GameMotivation_mean_new) + Gender + Age + Gender*Mvall + Gender*SOI + Gender*LeadershipDom + Gender*WordSum + Gender*Sexuality + Gender*RelStat + Gender*Income + Gender*Education + Gender*log(GameMotivation_mean_new), data=mturk2)

anova(VGplayself.all2.mturk2, test="Chisq")
summary(VGplayself.all2.mturk2)

etasq(VGplayself.all2.mturk2)
```



these are standardized coefficients...apparantly

```{r}
lm.beta(VGplayself.all.mturk2)
```

Plot Mate Value and VVGplay for each gender

```{r}
plotMV2 <- ggplot(mturk2, aes(x = Mvall, y = PlayXViolence_mean, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size=4) +
  ylim(0,50) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0.5,7.5), breaks=c(1:7)) +
  theme_bw() + geom_smooth(method=lm, fullrange=T) + 
  xlab("Mate Value") +
  ylab(NULL) +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position=c(.90,.10)) 

plotMV2
```

Now, plot both MV figures together
```{r}
grid.arrange(plotMV1, plotMV2, ncol=2)
g <- arrangeGrob(plotMV1, plotMV2, ncol=2)
ggsave(file="MateValue.tiff",g)
```

Now plot SOI
```{r}
plotSOI2 <- ggplot(mturk2, aes(x = SOI, y = PlayXViolence_mean, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size=4) +
  ylim(0,50) +
  scale_colour_hue(l=50, guide = FALSE) +
  scale_x_continuous(limits=c(0,80)) +
  theme_bw() + geom_smooth(method=lm,  fullrange=T) + 
  xlab("Sexual Interest") +
  ylab(NULL) +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position=c(.90,.10)) 
plotSOI2
```

Boxplot of VVGplay with education  (Self-reported)

```{r}
BoxEDU2 <- ggplot(mturk2, aes(x=Education, y=PlayXViolence_mean)) + geom_boxplot() + 
  guides(fill=FALSE) + 
  xlab("Level of Education") +
  ylab(NULL) +
  scale_x_discrete(labels=c("Primary", "High School", "Vocational", "University", "Post-Graduate"))+
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 

BoxEDU2
```

Now, plot both figures together
```{r}
grid.arrange(plotSOI1, plotSOI2, ncol=2)
g <- arrangeGrob(plotSOI1, plotSOI2, ncol=2)
ggsave(file="SOI.eps",g)
```

Now, plot both histograms of violent videogame play together
```{r}
grid.arrange(hist1, hist2, ncol=2)
g <- arrangeGrob(hist1, hist2, ncol=2)
ggsave(file="histogram.jpg",g)
```

And now both boxplots of education together
```{r}
grid.arrange(BoxEDU1, BoxEDU2, ncol=2)
g <- arrangeGrob(BoxEDU1, BoxEDU2, ncol=2)
ggsave(file="Education.tiff",g)
```

Now plot game_motivation
```{r}
ggplot(mturk2, aes(x = GameMotivation_mean_new, y = PlayXViolence_mean, color = Gender, shape = Gender, linetype = Gender)) + 
  geom_point(size=4) +
  theme_bw() + geom_smooth(method=lm, fullrange=T) + 
  scale_colour_hue(l=50, guide = FALSE) +
  ylim(0,50) +
  scale_x_continuous(limits=c(0.75,7.25), breaks=c(1:7)) +
  xlab("Gaming Motivation") +
  ylab("Violent Video Game Exposure") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position=c(.90,.10)) 

```


Next, I'm going to look at Kids, NumberGirls, NumberBoys, NumberTeenageBoys, KidsAge_mean, TeenageBoys_Binary.

but first I'll cut out the individuals that were never married as those individuals skew things at zero.

```{r}
mturk1Kids = subset(mturk1, RelStat %in% c("1", "3"))

VGplayself.family.mturk1<-lm(PlayXViolence_mean ~ NumberBoys + NumberGirls + Gender + Gender*NumberBoys + Gender*NumberGirls, data=mturk1Kids)
anova(VGplayself.family.mturk1, test="Chisq")
```

Looks like the number of boys a woman and man have affects how much violent video game exposure they have.

Let's plot that shit!!!

```{r}
ggplot(mturk1Kids, aes(x = NumberBoys, y = PlayXViolence_mean, color = Gender)) + geom_point(shape=1, size=4) +
  theme_bw() + geom_smooth(method=lm, se=FALSE, fullrange=T) + 
  xlab("The number of male offspring") +
  ylab("Violent Video Game Exposure") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
```


Supplementary Materials
-----------------------

Here is the full model with age and all the interactions for Mturk 1

```{r}
VGplayself.all.mturk1<-lm(PlayXViolence_mean ~ MV1.new + SOI + LeadershipDom + WordSum + CTSlog + Sexuality + RelStat + Income + Education + Gender + Age + Gender*MV1.new + Gender*SOI + Gender*LeadershipDom + Gender*WordSum + Gender*CTSlog + Gender*Sexuality + Gender*RelStat + Gender*Income + Gender*Education + Age*MV1.new + Age*SOI + Age*LeadershipDom + Age*WordSum + Age*CTSlog + Age*Sexuality + Age*RelStat + Age*Income + Age*Education, data=mturk1)

anova(VGplayself.all.mturk1, test="Chisq")
summary(VGplayself.all.mturk1)
```

and now for Mturk 2

```{r}
VGplayself.all.mturk2<-lm(PlayXViolence_mean ~ Mvall + SOI + WordSum + Sexuality + RelStat + Income + Education + Gender + Age + Gender*Mvall + Gender*SOI + Gender*WordSum + Gender*Sexuality + Gender*RelStat + Gender*Income + Gender*Education + Age*Mvall + Age*SOI + Age*WordSum + Age*Sexuality + Age*RelStat + Age*Income + Age*Education, data=mturk2)

anova(VGplayself.all.mturk2, test="Chisq")
summary(VGplayself.all.mturk2)
```


ADD ESRB RATING ANALYSES HERE AS EXTRA ANALYSES


